NanoMaLy

### 2015-11-18 paper/pldi16-submission/evaluation-nanomaly.tex
tweak
comment
trace size graph for seminal
move some text to the overview

\section{Evaluation: \nanomaly}
\label{sec:evaluation-nanomaly}
We have evaluated our prototype of \nanomaly in two phases:
%
\begin{enumerate}
\item We investigate the \emph{complexity} of the generated traces by various size metrics.
\item We investigate the \emph{usefulness} of the generated traces compared to the standard compiler errors by performing an A/B study.
\end{enumerate}

\subsection{Trace Complexity}
\label{sec:trace-complexity}
For each of the ill-typed programs from \S~\ref{sec:eval-witness} for
which we could find a witness, we measure the complexity of the
generated trace according to the following metrics:
%
\begin{enumerate}
\item The number of single-step reductions along the path from the
  witness to the stuck term. This can be thought of as a worst-case
  complexity, \ie ``How big is the fully-expanded trace?''
\item The number forward (or backward) jumps that can be taken along the
  path from the witness to the stuck term. These abstract many details
  of the computation that are often uninteresting or irrelevant to the
  explanation.
\item \ES{others?}
\end{enumerate}
%
\begin{figure}[t]
\centering
Have a graph!  
\caption{Complexity of the generated traces.}
\label{fig:results-complexity}
\end{figure}
%
The results of the experiment are summarized in
Figure~\ref{fig:results-complexity}.
%
The average number of single-step reductions is XX with a minimum of XX
and a maximum of XX. The average number of jumps is YY with a minimum of
YY and a maximium of YY.

>>>>

\section{Evaluation: \nanomaly}
\label{sec:evaluation-nanomaly}
We have implemented our interactive semantics for the pure subset of
\ocaml with a web-based visualization of the generated traces in a tool
called \nanomaly.
%
It takes as input an \ocaml program and a function to check, searches
for a witness as described in \S~\ref{sec:searching-witness}, and
produces as output an interactive trace of the (possibly buggy)
execution~\footnote{If no witness could be found we simply return a trace
  of the last attempt.}.

We have not really (\ES{huh?}) evaluated our prototype of \nanomaly in two phases:
%
\begin{enumerate}
\item We investigate the \emph{complexity} of the generated traces by various size metrics.
\item We investigate the \emph{usefulness} of the generated traces compared to the standard compiler errors by performing an A/B study.
\end{enumerate}

\subsection{Trace Complexity}
\label{sec:trace-complexity}
For each of the ill-typed programs from \S~\ref{sec:eval-witness} for
which we could find a witness, we measure the complexity of the
generated trace according to the following metrics:
%
\begin{enumerate}
\item The number of single-step reductions along the path from the
  witness to the stuck term. This can be thought of as a worst-case
  complexity, \ie ``How big is the fully-expanded trace?''
\item The number forward (or backward) jumps that can be taken along the
  path from the witness to the stuck term. These abstract many details
  of the computation that are often uninteresting or irrelevant to the
  explanation.
\item \ES{others?}
\end{enumerate}
%
\begin{figure*}[t]
\centering
\begin{minipage}{0.49\linewidth}
\includegraphics[width=\linewidth]{trace_size_seminal.pdf} % TODO: switch to ucsd
\end{minipage}
\begin{minipage}{0.49\linewidth}
\includegraphics[width=\linewidth]{trace_size_seminal.pdf}
\end{minipage}
\caption{Complexity of the generated traces.}
\label{fig:results-complexity}
\end{figure*}
%
The results of the experiment are summarized in
Figure~\ref{fig:results-complexity}.
%
The average number of single-step reductions is XX with a minimum of XX
and a maximum of XX. The average number of jumps is YY with a minimum of
YY and a maximium of YY.

### 2015-11-18 paper/pldi16-submission/overview.tex
move some text to the overview

>>>>

Our goal is to find concrete values that demonstrate how a
program ``goes wrong''.

\paragraph{Problem: Which inputs are bad?}
%
One approach is to randomly generate input values and use
them to execute the program until we find one that causes
the program to go wrong. However, to see why this approach
is naive, consider the following example:
%
\begin{code}
  let f x =
    let y = 1 + x in
      1. +. y
\end{code}
%
What \emph{types} of inputs should we test \texttt{f} with?
Values of type \texttt{int} and \texttt{float} are fair game,
but values of type say, \texttt{string} or \texttt{int list}
will cause the program to go wrong in an \emph{irrelevant}
manner.

\paragraph{Solution:} \RJ{STOP}

\subsection{Visualizing Witnesses}
\label{sec:visual-witness}
% We have implemented our interactive semantics for the pure subset of
% \ocaml with a web-based visualization of the generated traces in a tool
% called \nanomaly.
% %
% It takes as input an \ocaml program and a function to check, searches
% for a witness as described in \S~\ref{sec:searching-witness}, and
% produces as output an interactive trace of the (possibly buggy)
% execution~\footnote{If no witness could be found we simply return a trace
%   of the last attempt.}.
% %
We initially display a big-step reduction, \ie the source term followed
by the stuck term (or the final value).
%
The user can then progressively fill in the hidden steps of the
computation by selecting a visible term and choosing one of the
applicable traversal strategies -- described in
\S~\ref{sec:traversing-graph} -- to insert another term into the
visualization.

A sample interaction of the student program in
Figure~\ref{fig:palindrome} can be seen in
Figure~\ref{fig:nanomaly-palindrome}.
%
\begin{figure}[ht]
\centering
\begin{code}
  let listReverse l =
    let rec helper xs = function
      | []     -> xs
      | hd::tl -> helper (hd::xs) tl
    in helper []

  let palindrome w =
    listReverse w = w
\end{code}
\caption{Ill-typed \texttt{palindrome} function. The error lies in the
  use of the \texttt{function} construct which introduces an extra,
  undesired lambda abstraction, but manifests itself in the equality
  test on line XX. The fix is to replace \texttt{function} with
  \texttt{match xs with}. \ES{I think we should use this example in the
    intro/overview, leaving it here for now until they exist..}}
\label{fig:palindrome}
\end{figure}
%
\begin{figure}[ht]
% \centering
% Initial state: \\
% \includegraphics[width=\linewidth]{palindrome} \\
% After ``jumping backward'' from the stuck term: \\
% \includegraphics[width=\linewidth]{palindrome2} \\
% After ``stepping forward'' twice from @listReverse []@: \\
% \includegraphics[width=\linewidth]{palindrome3} \\
\caption{Sample interaction}
\label{fig:nanomaly-palindrome}
\end{figure}
%
The initial state of the visualization tells us that the witness is
@palindrome []@, and that after some number of steps -- the thick arrow
denotes a multi-step transition -- we try to compare a function to a
list, which is not allowed in \ocaml.
%
Based on our semantics for \lang one would expect the final state to be
the \stuck term, but in practice we simply halt execution at the point
where an expression would transition to \stuck, as this final state is
more informative.
%
Note also that the equality test is colored black while the contextual
if-expression is faded out, this tells the user that the program got
stuck while trying to evaluate the sub-expression.

Upon seeing the stuck term, one might wonder where the @function@
came from.
%
To investigate we select the stuck term and click the ``jump backward''
button to search backwards from the stuck term for the most recent
function call, which brings us to @listReverse [] = w@, in the same
context as before.
%
Uncontent with the explanation so far, we ``step forward'' twice from
the @listReverse []@ term, bringing us to @helper [] = w@.
%
At this point it is clear that the @helper@ function is not defined
correctly, we have supplied it with the single argument we expected and
yet it still returned a @function@.

The problem is that the @function@ keyword in \ocaml defines an
anonymous function that takes a single argument and immediately does a
case-analysis without giving the argument a name.
%
The solution is to replace @function@ with an explicit @match xs with@
-- naming the value we wish to case-analyse.
%
After applying our fix, \nanomaly -- and more importantly \ocaml --
decide that @listReverse@ is safe to run.
\ES{these last few paragraphs probably belong in the overview}
%

### 2015-11-18 paper/pldi16-submission/semantics.tex
move some text to the overview

Our goal is to find concrete values that demonstrate how a
program ``goes wrong''.

\paragraph{Problem: Which inputs are bad?}
%
One approach is to randomly generate input values and use
them to execute the program until we find one that causes
the program to go wrong. However, to see why this approach
is naive, consider the following example:
%
\begin{lstlisting}
  let f x =
    let y = 1 + x in
      1. +. y
\end{lstlisting}
%
What \emph{types} of inputs should we test \texttt{f} with?
Values of type \texttt{int} and \texttt{float} are fair game,
but values of type say, \texttt{string} or \texttt{int list}
will cause the program to go wrong in an \emph{irrelevant}
manner.

\paragraph{Solution:} \RJ{STOP}

>>>>

% Our goal is to find concrete values that demonstrate how a
% program ``goes wrong''.

% \paragraph{Problem: Which inputs are bad?}
% %
% One approach is to randomly generate input values and use
% them to execute the program until we find one that causes
% the program to go wrong. However, to see why this approach
% is naive, consider the following example:
% %
% \begin{lstlisting}
%   let f x =
%     let y = 1 + x in
%       1. +. y
% \end{lstlisting}
% %
% What \emph{types} of inputs should we test \texttt{f} with?
% Values of type \texttt{int} and \texttt{float} are fair game,
% but values of type say, \texttt{string} or \texttt{int list}
% will cause the program to go wrong in an \emph{irrelevant}
% manner.

% \paragraph{Solution:} \RJ{STOP}
